On the propriety of restricted Boltzmann machines

### 2016-11-21 writing/appendix/appendix-center.Rmd
nordman edits and references cleanup

---
title: "Supplementary Material to:  On the propriety of restricted Boltzmann machines"
author: "Andee Kaplan, Daniel Nordman, and Stephen Vardeman"
date: "November, 2016"
linestretch: 2 
output: 
  pdf_document:
    fig_caption: yes
    includes:
      in_header: ../latex/header.tex
    number_sections: yes
---

>>>>

---
title: "Supplementary material to:  On the propriety of restricted Boltzmann machines"
author: | 
  | Andee Kaplan, Daniel Nordman, and Stephen Vardeman
  | Iowa State University
  | ajkaplan@iastate.edu, dnordman@iastate.edu, vardeman@iastate.edu
date: "November, 2016"
linestretch: 2 
output: 
  pdf_document:
    fig_caption: yes
    includes:
      in_header: ../latex/header.tex
    number_sections: yes
---

### 2016-11-21 writing/appendix/appendix-note.Rmd
nordman edits and references cleanup

{\bf Proof of Proposition \ref{prop:instab}.} We prove the contrapositive. Suppose that $\Delta(\boldsymbol \theta_N) \le C$ holds for some $C > 0$. Under the FSFS model, $P_{\boldsymbol \theta_N}(\boldsymbol x) > 0$ holds for each outcome $\boldsymbol x \in \mathcal{X}^N$. Let $\boldsymbol x_{min} \equiv \argmin\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)$ and $\boldsymbol x_{max} \equiv \argmax\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)$. Note there exists a sequence $\boldsymbol x_{min} \equiv \boldsymbol x_0, \boldsymbol x_1, \dots, \boldsymbol x_k \equiv \boldsymbol x_{max}$ in $\mathcal{X}^N$ of component-wise switches to move from $\boldsymbol x_{min}$ to $\boldsymbol x_{max}$ in the sample space (i.e. $\boldsymbol x_i, \boldsymbol x_{i + 1} \in \mathcal{X}^N$ differ by exactly $1$ component for $i = 0, \dots, k$) for some integer $k \in \{0, 1, \dots, N\}$. Then, if $k > 0$, it follows that
\begin{align*}
\log\left[\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_{max})}{P_{\boldsymbol \theta_N}(\boldsymbol x_{min})}\right] &= \left|\sum\limits_{i = 1}^k\log\left(\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_i)}{P_{\boldsymbol \theta_N}(\boldsymbol x_{i-1})}\right)\right| \le \sum\limits_{i = 1}^k\left|\log\left(\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_i)}{P_{\boldsymbol \theta}(\boldsymbol x_{i-1})}\right)\right| \le k \Delta_N(\boldsymbol \theta_N) \le NC
\end{align*}
using $k \le N$ and $\Delta(\boldsymbol \theta_N) \le C$. If $k = 0$, then $\boldsymbol x_{max} = \boldsymbol x_{min}$ and the same bound above holds.$\Box$

{\bf Proof of Proposition \ref{prop:degen}.} Define $\boldsymbol x_{max}$ and $\boldsymbol x_{min}$ as in the proof of Proposition \ref{prop:instab}. Fix $0 < \epsilon < 1$ Then, $\boldsymbol x_{max} \in M_{\epsilon, \boldsymbol \theta_N}$, so $P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N}) \ge P_{\boldsymbol \theta_N}(\boldsymbol x_{max})$. If $\boldsymbol x \in \mathcal{X}^N \setminus M_{\epsilon, \boldsymbol \theta_N}$, then by definition $P_{\boldsymbol \theta_N}(\boldsymbol x) \le [P_{\boldsymbol \theta_N}(\boldsymbol x_{max})]^{1-\epsilon}[P_{\boldsymbol \theta_N}(\boldsymbol x_{min})]^{\epsilon}$ holds so that
\begin{align*}
1-P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})
& = \sum\limits_{\boldsymbol x \in \mathcal{X}^N \setminus M_{\epsilon, \boldsymbol \theta_N}}P_{\boldsymbol \theta_N}(\boldsymbol x) \\
& \le (2^N)[P_{\boldsymbol \theta_N}(\boldsymbol x_{max})]^{1-\epsilon}[P_{\boldsymbol \theta_N}(\boldsymbol x_{min})]^{\epsilon}
\end{align*}
Then, it follows that
\begin{align*}
\frac{1}{N}\log\left[\frac{P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})}{1-P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})}\right] & \ge \frac{1}{N} \log\left[\frac{P_{\boldsymbol\theta_N}(\boldsymbol x_{max})}{(2^V)[P_{\boldsymbol \theta_N}(\boldsymbol x_{max})]^{1-\epsilon}[P_{\boldsymbol \theta_N}(\boldsymbol x_{min})]^{\epsilon}}\right] \\
&= \frac{\epsilon}{N} \log\left[\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_{max})}{P_{\boldsymbol \theta_N}(\boldsymbol x_{min})}\right] - \log 2 \rightarrow \infty
\end{align*}
as $N \rightarrow \infty$ by the definition of an unstable FSFS model. Consequently, $P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N}) \rightarrow 1$ as $N \rightarrow \infty$ as claimed.$\Box$

>>>>

{\bf Proof of Proposition \ref{prop:instab}.} We prove the contrapositive, supposing that $\Delta(\boldsymbol \theta_N) \le C$ holds for some $C > 0$ and show $\text{ELPR}(\boldsymbol \theta_N) \leq C$. Let $\boldsymbol x_{min} \equiv \argmin\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)$ and $\boldsymbol x_{max} \equiv \argmax\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)$. Note there exists a sequence $\boldsymbol x_{min} \equiv \boldsymbol x_0, \boldsymbol x_1, \dots, \boldsymbol x_k \equiv \boldsymbol x_{max}$ in $\mathcal{X}^N$ of component-wise switches to move from $\boldsymbol x_{min}$ to $\boldsymbol x_{max}$ in the sample space (i.e. $\boldsymbol x_i, \boldsymbol x_{i + 1} \in \mathcal{X}^N$ differ by exactly $1$ component for $i = 0, \dots, k$) for some integer $k \in \{0, 1, \dots, N\}$. Under the FSFS model, recall $P_{\boldsymbol \theta_N}(\boldsymbol x) > 0$ holds so that $\log P_{\boldsymbol \theta_N}(\boldsymbol x)$ is well-defined for each outcome $\boldsymbol x \in \mathcal{X}^N$. Then, if $k > 0$, it follows that
$$
N\cdot\text{ELPR}(\boldsymbol \theta_N) = \log\left[\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_{max})}{P_{\boldsymbol \theta_N}(\boldsymbol x_{min})}\right] = \left|\sum\limits_{i = 1}^k\log\left(\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_i)}{P_{\boldsymbol \theta_N}(\boldsymbol x_{i-1})}\right)\right| \le \sum\limits_{i = 1}^k\left|\log\left(\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_i)}{P_{\boldsymbol \theta}(\boldsymbol x_{i-1})}\right)\right| \le k \Delta_N(\boldsymbol \theta_N) \le NC,
$$
using $k \le N$ and $\Delta(\boldsymbol \theta_N) \le C$. If $k = 0$, then $\boldsymbol x_{max} = \boldsymbol x_{min}$ and the same bound above holds.$\Box$

{\bf Proof of Proposition \ref{prop:degen}.} Define $\boldsymbol x_{max}$ and $\boldsymbol x_{min} \in \mathcal{X}^N$ as in the proof of Proposition \ref{prop:instab} where $|\mathcal{X}|<\infty$ holds in the FSFS model. We may suppose $|\mathcal{X}|>1$ (i.e., $\mathcal{X}^N$ has more than one outcome) because otherwise the model is trivially degenerate for all $N \geq 1$. Fix $0 < \epsilon < 1$ Then, $\boldsymbol x_{max} \in M_{\epsilon, \boldsymbol \theta_N}$, so $P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N}) \ge P_{\boldsymbol \theta_N}(\boldsymbol x_{max}) > 0$. If $\boldsymbol x \in \mathcal{X}^N \setminus M_{\epsilon, \boldsymbol \theta_N}$, then by definition $P_{\boldsymbol \theta_N}(\boldsymbol x) \le [P_{\boldsymbol \theta_N}(\boldsymbol x_{max})]^{1-\epsilon}[P_{\boldsymbol \theta_N}(\boldsymbol x_{min})]^{\epsilon}$ holds so that
\begin{align*}
1-P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})
& = \sum\limits_{\boldsymbol x \in \mathcal{X}^N \setminus M_{\epsilon, \boldsymbol \theta_N}}P_{\boldsymbol \theta_N}(\boldsymbol x) \\
& \le (|\mathcal{X}|^N)[P_{\boldsymbol \theta_N}(\boldsymbol x_{max})]^{1-\epsilon}[P_{\boldsymbol \theta_N}(\boldsymbol x_{min})]^{\epsilon}.
\end{align*}
From the lower bound on $P_{\boldsymbol \theta_N}(\boldsymbol x_{max})$ and the upper bound on $1-P_{\boldsymbol \theta_N}(\boldsymbol x_{max})$, it follows that
\begin{align*}
\frac{1}{N}\log\left[\frac{P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})}{1-P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})}\right] & \ge \frac{1}{N} \log\left[\frac{P_{\boldsymbol\theta_N}(\boldsymbol x_{max})}{(|\mathcal{X}|^N)[P_{\boldsymbol \theta_N}(\boldsymbol x_{max})]^{1-\epsilon}[P_{\boldsymbol \theta_N}(\boldsymbol x_{min})]^{\epsilon}}\right] \\
&= \frac{\epsilon}{N} \log\left[\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_{max})}{P_{\boldsymbol \theta_N}(\boldsymbol x_{min})}\right] - \log |\mathcal{X}| \rightarrow \infty
\end{align*}
as $N \rightarrow \infty$ by the definition of an unstable FSFS model (c.f. Definition \ref{defn:instab}). Consequently, $P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N}) \rightarrow 1$ as $N \rightarrow \infty$ as claimed.$\Box$

### 2016-11-21 writing/draft.Rmd
nordman edits and references cleanup

---
title: "On the propriety of restricted Boltzmann machines"
author: "Andee Kaplan, Daniel Nordman, and Stephen Vardeman"
date: "November, 2016"
linestretch: 2 
output: 
  pdf_document:
    fig_caption: yes
    includes:
      in_header: latex/header.tex
    number_sections: yes
bibliography: references/refs.bib
abstract: |
  A restricted Boltzmann machine (RBM) is an undirected graphical model constructed for discrete or continuous random variables, with two layers, one hidden and one visible, and no conditional dependency within a layer. In recent years, RBMs have risen to prominence due to their connection to deep learning. By treating a hidden layer of one RBM as the visible layer in a second RBM, a deep architecture can be created. RBMs are thought to thereby have the ability to encode very complex and rich structures in data, making them attractive for supervised learning. However, the generative behavior of RBMs is largely unexplored. In this paper, we discuss the relationship between RBM parameter specification in the binary case and the prevalence of undesirable model properties such as degeneracy, instability and uninterpretability. We also describe the difficulties that arise in likelihood-based and Bayes fitting  of such (highly flexible) models, especially as Gibbs sampling (quasi-Bayes) methods are often advocated for the RBM model structure.
---

@@

The data mining and machine learning communities have recently shown great interest in deep learning, specifically in stacked restricted Boltzmann machines (RBMs) [see @salakhutdinov2009deep, @salakhutdinov2012efficient, @srivastava2013modeling, @le2008representational for examples]. A RBM is a probabilistic undirected graphical model (for discrete or continuous random variables) with two layers, one hidden and one visible, with no conditional dependency within a layer [@smolensky1986information]. These models have reportedly been used with success in classification of images [@larochelle2008classification, @srivastava2012multimodal]. However, the model properties are largely unexplored in the literature and the commonly cited fitting methodology remains heuristic-based and abstruse [@hinton2006fast]. In this paper, we provide steps toward a thorough understanding of the model class and its properties from the perspective of statistical theory, and then explore the possibility of a rigorous fitting methodology. We find the RBM model class to be deficient in two fundamental ways.

First, the models are often unsatisfactory as conceptualizations of how data are generated. Recalling [@fisher1922mathematical], the aim of a statistical model is to represent data in a compact way. Neyman and Box further state that a model should "provide an explanation of the mechanism underlying the observed phenomena" [@lehmann1990model, @box1967discrimination]. At issue, RBMs  often produce data lacking realistic variability and may thereby fail to provide a satisfactory conceptualization of a data generation process. In addition to such degeneracy, we find that RBMs can easily exhibit types of instability. In practice, this may seen when a single pixel change in an image results in a wildly different classification in an image classification problem. Such model impropriety issues have recently been documented in RBMs [@li2014biclustering], as well as other deep architectures [@szegedy2013intriguing, @nguyen2014deep]. We investigate these phenomena for RBMs in Section \ref{explorations} through simulations of small, manageable examples.

Secondly, the fitting of these models is problematic. As the size of these models grows, both maximum likelihood and Bayesian methods of fitting quickly become intractable. The literature often suggests Markov chain Monte Carlo (MCMC) tools for approximate maximization of likelihoods to fit these models (e.g., Gibbs sampling to exploit conditional structure in hidden and visible variables), but little is said about acheiving convergence [@hinton2010practical, @hinton2006fast]. Related to this, these MCMC algorithms require updating potentially many latent variables (hiddens) which can critically influence convergence in MCMC-based likelihood methods.

@@

For a RBM model with enough hidden variables, it has been shown that any distribution for the visibles can be approximated arbitrarily well [@le2008representational, @montufar2011refinements, and @montufar2011expressive]. However, the empirical distribution of a training set of vectors of visibles is the best fitting model for observed cell data. As a consequence, we find that any fully principled fitting method based on the likelihood for RBM with enough hidden variables will simply seek to reproduce the (discrete) empirical distribution of a training set. Thus, there is no "smoothed distribution" achieved in fitting a RBM model of sufficient size with a rigorous likelihood-based method. We are therefore led to be skeptical that any model built using these structures (like a deep Boltzmann machine) can achieve prediction or inference in a principled way without intentially limiting the flexibility of the fitted model.

@@

RBMs have risen to prominence in recent years due to their connection to deep learning [see @hinton2006fast, @salakhutdinov2012efficient, or @srivastava2013modeling for examples]. By stacking multiple layers of RBMs in a deep architecture, proponents of the models claim to produce the ability to learn "internal representations that become increasingly complex, which is considered to be a promising way of solving object and speech recognition problems" [@salakhutdinov2009deep, pp. 450]. The stacking is achieved by treating a hidden layer of one RBM as the visible layer in a second RBM, and so on, until the desired multi-layer architecture is created.

@@

The previous discussion intends to indicate that the RBM model class suffers from parameter identifiability issues that go beyond mere symmetries in the parametrization. Not only it is possible to approximate any distribution on the visibles arbitrarily well (cf. @montufar2011refinements), quite different parameter settings can induce the same essential RBM model. However, this is not the most disastrous implication of the RBM parameterization. A far worse consequence is that, when fitting the RBM model by likelihood-based methods, we already know the nature of the answer before we begin: namely, such  fitting will simply aim to reproduce the empirical distribution from the training data if sufficiently many hiddens are in the model. That is, based on a random sample of vectors of visible variables, the model for the cell probabilities that has the highest likelihood over  *all possible model classes* (i.e., RBM-based or not) is the empirical distribution, and the parametrization of the RBM model itself ensures that this empirical distribution can be arbitrarily well-approximated.

For illustration, continue the simple example from above with $n$ iid observations, each consisting of two realized visibles $(V_1,V_2)$.  In which case, when the specified cell probabilities $p_{(-1,-1)},p_{(1,-1)}, p_{(-1,1)},p_{(1,1)}$ in (\ref{eqn:0}) are taken as the empirical cell frequencies from the sample, there is no better model based on maximum likelihood and the discussion above (c.f. (\ref{eqn:2})) shows that RBM model parameters can be chosen to re-create this empirical distribution to an arbitrarily close degree. Hence, RBM model fitting based on ML will simply seek to reproduce the empirical distribution. What's more, whenever this empirical distribution contains empty cells, fitting steps for the RBM model will necessarily aim to choose parameters that necessarily diverge to infinity in magnitude in order to zero-out the corresponding RBM cell probabilities. In data applications with a large sample space, it is unlikely that the training set will include at least one of each possible vector outcome (unlike this small example). This implies that some RBM model parameters must diverge to $+\infty$ to mimic the empirical distribution with empty cells and, as we have already discussed in Section \ref{explorations}, large magnitudes of $\boldsymbol \theta$ lead to model impropriety in the RBM.

@@

\begin{table}[ht]
\centering
\caption{The values used for the hyperparameters for all three fitting methods. A rule of thumb is imposed which decreases prior variances for the model parameters as the size of the model increases and also shrinks $\boldsymbol \theta_{interaction}$ more than $\boldsymbol \theta_{main}$.}
\label{tab:hyperparam}
\begin{tabular}{|l|c|c|}
\hline 
Method & Hyperparameter & Value \\ 
\hline \hline
\multirow{2}{*}{BwTPLV} & $C_1$ & $\frac{C}{n}\frac{1}{\nh + \nv}$ \\
 & $C_2$ & $\frac{C}{n}\frac{1}{\nh*\nv}$ \\
\hline
\multirow{2}{*}{BwTNLV} & $\sigma^2_{main}$ & $\frac{1}{\nh + \nv}$ \\
 & $\sigma^2_{interaction}$ & $\frac{1}{\nh*\nv}$ \\
\hline
\multirow{2}{*}{BwTNML} & $\sigma^2_{main}$ & $\frac{1}{\nh + \nv}$ \\
 & $\sigma^2_{interaction}$ & $\frac{1}{\nh*\nv}$ \\
\hline
\end{tabular}
\end{table}

>>>>

---
title: "On the propriety of restricted Boltzmann machines"
author: | 
  | Andee Kaplan, Daniel Nordman, and Stephen Vardeman
  | Iowa State University
  | ajkaplan@iastate.edu, dnordman@iastate.edu, vardeman@iastate.edu
date: "November, 2016"
linestretch: 2 
output: 
  pdf_document:
    fig_caption: yes
    includes:
      in_header: latex/header.tex
    number_sections: yes
bibliography: references/refs.bib
abstract: |
  A restricted Boltzmann machine (RBM) is an undirected graphical model constructed for discrete or continuous random variables, with two layers, one hidden and one visible, and no conditional dependency within a layer. In recent years, RBMs have risen to prominence due to their connection to deep learning. By treating a hidden layer of one RBM as the visible layer in a second RBM, a deep architecture can be created. RBMs are thought to thereby have the ability to encode very complex and rich structures in data, making them attractive for supervised learning. However, the generative behavior of RBMs is largely unexplored. In this paper, we discuss the relationship between RBM parameter specification in the binary case and the prevalence of undesirable model properties such as degeneracy, instability and uninterpretability. We also describe the difficulties that arise in likelihood-based and Bayes fitting of such (highly flexible) models, especially as Gibbs sampling (quasi-Bayes) methods are often advocated for the RBM model structure.
---

@@

The data mining and machine learning communities have recently shown great interest in deep learning, specifically in stacked restricted Boltzmann machines (RBMs) [see @salakhutdinov2009deep; @salakhutdinov2012efficient; @srivastava2013modeling; @le2008representational for examples]. A RBM is a probabilistic undirected graphical model (for discrete or continuous random variables) with two layers, one hidden and one visible, with no conditional dependency within a layer [@smolensky1986information]. These models have reportedly been used with success in classification of images [@larochelle2008classification; @srivastava2012multimodal]. However, the model properties are largely unexplored in the literature and the commonly cited fitting methodology remains heuristic-based and abstruse [@hinton2006fast]. In this paper, we provide steps toward a thorough understanding of the model class and its properties from the perspective of statistical theory, and then explore the possibility of a rigorous fitting methodology. We find the RBM model class to be deficient in two fundamental ways.

First, the models are often unsatisfactory as conceptualizations of how data are generated. Recalling [@fisher1922mathematical], the aim of a statistical model is to represent data in a compact way. Neyman and Box further state that a model should "provide an explanation of the mechanism underlying the observed phenomena" [@lehmann1990model; @box1967discrimination]. At issue, RBMs  often produce data lacking realistic variability and may thereby fail to provide a satisfactory conceptualization of a data generation process. In addition to such degeneracy, we find that RBMs can easily exhibit types of instability. In practice, this may seen when a single pixel change in an image results in a wildly different classification in an image classification problem. Such model impropriety issues have recently been documented in RBMs [@li2014biclustering], as well as other deep architectures [@szegedy2013intriguing; @nguyen2014deep]. We investigate these phenomena for RBMs in Section \ref{explorations} through simulations of small, manageable examples.

Secondly, the fitting of these models is problematic. As the size of these models grows, both maximum likelihood and Bayesian methods of fitting quickly become intractable. The literature often suggests Markov chain Monte Carlo (MCMC) tools for approximate maximization of likelihoods to fit these models (e.g., Gibbs sampling to exploit conditional structure in hidden and visible variables), but little is said about acheiving convergence [@hinton2010practical; @hinton2006fast]. Related to this, these MCMC algorithms require updating potentially many latent variables (hiddens) which can critically influence convergence in MCMC-based likelihood methods.

@@

For a RBM model with enough hidden variables, it has been shown that any distribution for the visibles can be approximated arbitrarily well [@le2008representational; @montufar2011refinements; and @montufar2011expressive]. However, the empirical distribution of a training set of vectors of visibles is the best fitting model for observed cell data. As a consequence, we find that any fully principled fitting method based on the likelihood for RBM with enough hidden variables will simply seek to reproduce the (discrete) empirical distribution of a training set. Thus, there is no "smoothed distribution" achieved in fitting a RBM model of sufficient size with a rigorous likelihood-based method. We are therefore led to be skeptical that any model built using these structures (like a deep Boltzmann machine) can achieve prediction or inference in a principled way without intentially limiting the flexibility of the fitted model.

@@

RBMs have risen to prominence in recent years due to their connection to deep learning [see @hinton2006fast; @salakhutdinov2012efficient; @srivastava2013modeling for examples]. By stacking multiple layers of RBMs in a deep architecture, proponents of the models claim to produce the ability to learn "internal representations that become increasingly complex, which is considered to be a promising way of solving object and speech recognition problems" [@salakhutdinov2009deep, pp. 450]. The stacking is achieved by treating a hidden layer of one RBM as the visible layer in a second RBM, and so on, until the desired multi-layer architecture is created.

@@

The previous discussion intends to indicate that the RBM model class suffers from parameter identifiability issues that go beyond mere symmetries in the parametrization. Not only it is possible to approximate any distribution on the visibles arbitrarily well [cf. @montufar2011refinements], quite different parameter settings can induce the same essential RBM model. However, this is not the most disastrous implication of the RBM parameterization. A far worse consequence is that, when fitting the RBM model by likelihood-based methods, we already know the nature of the answer before we begin: namely, such  fitting will simply aim to reproduce the empirical distribution from the training data if sufficiently many hiddens are in the model. That is, based on a random sample of vectors of visible variables, the model for the cell probabilities that has the highest likelihood over  *all possible model classes* (i.e., RBM-based or not) is the empirical distribution, and the parametrization of the RBM model itself ensures that this empirical distribution can be arbitrarily well-approximated.

For illustration, continue the simple example from above with $n$ iid observations, each consisting of two realized visibles $(V_1,V_2)$.  In which case, when the specified cell probabilities $p_{(-1,-1)},p_{(1,-1)}, p_{(-1,1)},p_{(1,1)}$ in (\ref{eqn:0}) are taken as the empirical cell frequencies from the sample, there is no better model based on maximum likelihood and the discussion above (cf. (\ref{eqn:2})) shows that RBM model parameters can be chosen to re-create this empirical distribution to an arbitrarily close degree. Hence, RBM model fitting based on ML will simply seek to reproduce the empirical distribution. What's more, whenever this empirical distribution contains empty cells, fitting steps for the RBM model will necessarily aim to choose parameters that necessarily diverge to infinity in magnitude in order to zero-out the corresponding RBM cell probabilities. In data applications with a large sample space, it is unlikely that the training set will include at least one of each possible vector outcome (unlike this small example). This implies that some RBM model parameters must diverge to $+\infty$ to mimic the empirical distribution with empty cells and, as we have already discussed in Section \ref{explorations}, large magnitudes of $\boldsymbol \theta$ lead to model impropriety in the RBM.

@@

\begin{table}[ht]
\centering
\caption{The values used for the hyperparameters for all three fitting methods. A rule of thumb is imposed which decreases prior variances for the model parameters as the size of the model increases and also shrinks $\boldsymbol \theta_{interaction}$ more than $\boldsymbol \theta_{main}$. The common $C$ defining $C_1$ and $C_2$  in the BwTPLV method is chosen by tuning.}
\label{tab:hyperparam}
\begin{tabular}{|l|c|c|}
\hline 
Method & Hyperparameter & Value \\ 
\hline \hline
\multirow{2}{*}{BwTPLV} & $C_1$ & $\frac{C}{n}\frac{1}{\nh + \nv}$ \\
 & $C_2$ & $\frac{C}{n}\frac{1}{\nh*\nv}$ \\
\hline
\multirow{2}{*}{BwTNLV} & $\sigma^2_{main}$ & $\frac{1}{\nh + \nv}$ \\
 & $\sigma^2_{interaction}$ & $\frac{1}{\nh*\nv}$ \\
\hline
\multirow{2}{*}{BwTNML} & $\sigma^2_{main}$ & $\frac{1}{\nh + \nv}$ \\
 & $\sigma^2_{interaction}$ & $\frac{1}{\nh*\nv}$ \\
\hline
\end{tabular}
\end{table}
